\section{Related Work}
\label{sec:related}

\para{Activation steering.}
\citet{turner2023steering} introduce Activation Addition (\actadd), which steers language models by adding the difference in activations between a pair of contrastive prompts to the residual stream. \citet{rimsky2024steering} scale this to Contrastive Activation Addition (\caa), averaging activation differences across hundreds of contrastive pairs and applying the resulting vector at every generated token position. \citet{li2023inference} propose Inference-Time Intervention (\iti), which shifts activations at a limited number of attention heads to improve truthfulness. All three methods apply steering continuously at every forward pass, implicitly assuming that one-shot steering would be insufficient---an assumption our work directly tests and confirms.

\para{Steering schedules and decay.}
\citet{scalena2024dynamic} compare multiple steering schedules: start-only (first step), fixed (constant $\alpha$), diminishing (linear decay), and dynamic (KL-divergence-adaptive). They show that start-only steering fails to maintain behavioral conditioning, while their dynamic method reveals that the optimal steering intensity drops sharply after the first few tokens. \citet{xie2025comparative} propose a token-wise decaying strategy $\alpha_t = \alpha \cdot (1/(1 + \omega t))^k$ for SAE-based steering, designed to prevent degeneration from constant high-strength intervention. \citet{klerings2025steering} systematically study steering strength, location, and duration for grammatical tense, finding that late or extended steering causes topic shift and degeneration. These works prescribe how steering \emph{should} decay but do not measure how it \emph{naturally} decays in hidden states after removal. Our work fills this gap with direct measurements.

\para{Representation geometry of steered behaviors.}
\citet{arditi2024refusal} show that refusal in language models is mediated by a single direction in the residual stream, demonstrating that some behaviors have strong, persistent linear representations. \citet{tan2024analyzing} evaluate the generalization and reliability of steering vectors, finding substantial variability across inputs and spurious biases. \citet{subramani2022extracting} demonstrate that steering vectors can precisely control generation when applied correctly. Our decay measurements complement this line of work by characterizing the \emph{temporal} persistence of steered representations across generation steps.

\para{KV cache steering.}
\citet{belitsky2025kv} propose one-shot steering via KV cache modification, where cached key-value pairs are modified once before generation and persist through the attention mechanism. They contrast this with activation steering, which requires continuous intervention, implicitly demonstrating that activation steering decays while KV cache modifications persist. Our results provide direct evidence for this contrast: we show that the KV cache is the primary vehicle for any residual steering signal after activation-based intervention stops, and that removing steered KV pairs causes the model to overcorrect.
